Do researchers preferentially collaborate with same-gendered colleagues?

### 2018-05-25 Gender homophilly project.Rmd
Added simulation

>>>>

## Theoretical expectations for $\alpha$ when there is a gender gap between career stages

This section makes Figure 6, by deriving the expected $\alpha$ under various assumptions, assuming that there is (potentially) a gender gap between career stages but no direct choice of collaborators by gender.

```{r}
simulate <- function(params){

  # Attach the vectors of paramters from the dataframe input
  GR.stage1 <- params$GR.stage1
  GR.stage2 <- params$GR.stage2
  prop.student.prof <- params$prop.student.prof
  prop.student.student <- params$prop.student.student

  # This is the proportion of each type of collaboration:
  # early-early, early-late, late-late respectively
  p1.1 <- (1 - prop.student.prof) * prop.student.student
  p1.2 <- prop.student.prof
  p2.2 <- (1 - prop.student.prof) * (1-prop.student.student)

  # given the gender ratio at career stages a and b, find the proportion of collaborations that are man-man, woman-woman, or mixed gender
  man.man <- function(GR_a, GR_b) (1-GR_a) * (1-GR_b)
  woman.woman <- function(GR_a, GR_b) GR_a * GR_b
  mixed <- function(GR_a, GR_b) (1-GR_a) * GR_b + (1-GR_b) * GR_a

  # Find the frequencies of the 3 kinds of collaboration
  man.man <- p1.1 * man.man(GR.stage1, GR.stage1) +
    p1.2 * man.man(GR.stage1, GR.stage2) +
    p2.2 * man.man(GR.stage2, GR.stage2)

  woman.woman <- p1.1 * woman.woman(GR.stage1, GR.stage1) +
    p1.2 * woman.woman(GR.stage1, GR.stage2) +
    p2.2 * woman.woman(GR.stage2, GR.stage2)

  mixed <- p1.1 * mixed(GR.stage1, GR.stage1) +
    p1.2 * mixed(GR.stage1, GR.stage2) +
    p2.2 * mixed(GR.stage2, GR.stage2)

  # Find p and q, which are used to calcualte alpha

  # p is proportion of men coauthoring with men
  p <- 1 - mixed / (2 * man.man + mixed)
  # q is proportion of women coauthoring with men
  q <- mixed / (2 * woman.woman + mixed)

  data.frame(alpha = p - q, 
             GR.stage1,
             GR.stage2,
             prop.student.prof,
             prop.student.student)
}

results <- expand.grid(
  GR.stage1 = seq(0.01, 0.7, length = 201),
  GR.stage2 = seq(0.01, 0.7, length = 201),
  prop.student.prof = c(0.01, 0.25, 0.5, 0.75, .99), 
  prop.student.student = c(0.01, 0.25, 0.5, 0.75, .99)) %>% 
  filter(GR.stage1 >= GR.stage2) %>%
  simulate() %>% 
  mutate(GR.stage1 = round(GR.stage1 * 100),
         GR.stage2 = round(GR.stage2 * 100),
         prop.student.prof = prop.student.prof * 100,
         prop.student.student = prop.student.student * 100) %>%
  filter(!duplicated(paste(GR.stage1,GR.stage2,prop.student.prof,prop.student.student)))

simulation.figure <- results %>%
  ggplot(aes(GR.stage1, GR.stage2, z = alpha, fill = alpha)) + 
  facet_grid(prop.student.prof ~ prop.student.student) + 
  geom_tile() + 
  #stat_contour(binwidth = 0.1, colour = "grey20", size = 0.3) + 
  scale_fill_distiller(name = "\u03B1",
                       palette = "RdYlBu",
                       limits = c(-1,1) * max(abs(results$alpha))) +
  scale_x_continuous(expand = c(0,0), limits = c(0,70))+
  scale_y_continuous(expand = c(0,0), limits = c(0,70))+
  xlab("% women among early-career authors") + 
  ylab("% women among established authors") +
  theme_bw() + 
  theme(panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        legend.position = "bottom") + 
  ggtitle("Rows show % between-stage collaborations ", 
          subtitle = "Columns show % within-stage collaborations that involve two early-career researchers")

simulation.figure
ggsave(simulation.figure, file = "figures/figure5.pdf", 
       height = 5.6, width = 5)#, device = cairo_pdf)
```

### R session info
```{r}
sessionInfo() %>% pander()
```
